19Q1

### 2019-04-29 19Q1.Rmd
updated Rmd

    # Get all G2Pgene-cell line combinations
    grid <- expand.grid("Cell_line" = unique(filt$Cell_line), "Gene" = g2p_genes)
    data <-  merge(filt, grid, by = "Cell_line", all.y = TRUE)
    print("Merged grid.")
    # Merge metadata
    data <-  merge(data, select(ccl_meta, Cell_line, Histology, CCLE_Name, Broad_ID, Primary.Disease), by = c("Cell_line", "Histology", "CCLE_Name", "Broad_ID"), all.x = TRUE)
    ## Add columns for G2P drug-gene and drug-gene-lineage sorting
    data$Drug_Gene <- paste0(data$Drug, "_", data$Gene)
    data$G2P_Drug_Gene <- ifelse(data$Drug_Gene %in% g2p_druggenes, "Yes", "No")
    data$Drug_Gene_Lin <- paste0(data$Drug, "_", data$Gene, "_", data$Primary.Disease)
    data$G2P_Drug_Gene_Lin <- ifelse(data$Drug_Gene_Lin %in% g2p_dgls, "Yes", "No")
    print("Merged metadata.")
  }
  if(database == "CRISPR") {
    print(paste0("Processing ", database, " data."))
    # Melt CRISPR dataset for merging
    data <- melt(crispr, id.vars = "Broad_ID", measure.vars = colnames(crispr)[2:ncol(crispr)], variable.name = "Gene", value.name = "Score")
    # Merge cell line metadata
    data <- merge(data, select(ccl_meta, Cell_line, Histology, CCLE_Name, Broad_ID, Primary.Disease), by = "Broad_ID", all.x = TRUE)
    data$Drug_Gene <- paste0(data$Drug, "_", data$Gene)
    data$Drug_Gene_Lin <- paste0(data$Drug, "_", data$Gene, "_", data$Primary.Disease)
    print("Merged metadata.")
  }

@@

ccle <- MergeOmicsG2P(database = "CCLE", outfile = "./data_munging/rds/ccle_data_19Q1_g2pgenes.rds")
# ctrp <- MergeOmicsG2P(database = "CTRP", outfile = "./../htscreens_giant_files/ctrp_data_19Q1_g2pgenes.rds")
# gdsc <- MergeOmicsG2P(database = "GDSC", outfile = "./data_munging/rds/gdsc_data_19Q1_g2pgenes.rds")

ccle_ptmuts <- FilterForPointMuts(dataframe = ccle, outfile = "./data_munging/rds/ccle_data_19Q1_g2pgenes_ptmuts.rds")
# ctrp_ptmuts <- FilterForPointMuts(dataframe = ctrp, outfile = "./data_munging/rds/ctrp_data_19Q1_g2pgenes_ptmuts.rds")
# gdsc_ptmuts <- FilterForPointMuts(dataframe = gdsc, outfile = "./data_munging/rds/gdsc_data_19Q1_g2pgenes_ptmuts.rds")
```

@@

```{r, eval=TRUE}
ccle <- readRDS("./data_munging/rds/ccle_data_19Q1_g2pgenes.rds")
# ctrp <- readRDS("./../htscreens_giant_files/ctrp_data_19Q1_g2pgenes.rds")
gdsc <- readRDS("./data_munging/rds/gdsc_data_19Q1_g2pgenes.rds")
ccle_ptmuts <- readRDS("./data_munging/rds/ccle_data_19Q1_g2pgenes_ptmuts.rds")
# ctrp_ptmuts <- readRDS("./../htscreens_giant_files/ctrp_data_19Q1_g2pgenes_ptmuts.rds")
gdsc_ptmuts <- readRDS("./data_munging/rds/gdsc_data_19Q1_g2pgenes_ptmuts.rds")

# Testing
test_ccle <- ccle_ptmuts %>% count(Gene, Mutation_Status)
# test_ctrp <- ctrp_ptmuts %>% count(Gene, Mutation_Status)
test_gdsc <- gdsc_ptmuts %>% count(Gene, Mutation_Status)
```

@@

```{r, eval=TRUE}
crispr_all <- readRDS("./../htscreens_giant_files/crispr_data_19Q1.rds")

# Filter G2P CRISPR for point mutations
crispr_ptmuts <- filter(crispr_data, PointMutation == 1)
crispr_ptmuts <- subset(crispr_data_ptmuts, select = -c(PointMutation))
# Filter CRISPR for G2P genes in G2P
crispr_data_g2p <- filter(crispr_ptmuts, Gene %in% g2p_genes)
crispr_data_g2p <- unique(crispr_data_g2p)
crispr_data_g2p$Gene <- factor(crispr_data_g2p$Gene)
crispr_data_g2p$Mutation_Status <- factor(crispr_data_g2p$Mutation_Status)
crispr_data_g2p$Primary.Disease <- factor(crispr_data_g2p$Primary.Disease)
```

Test merging and filtering functions on CRISPR data:
```{r, eval=FALSE}
crispr <- MergeOmicsG2P(database = "CRISPR", outfile = "./../htscreens_giant_files/crispr_data_19Q1_g2pgenes.rds")
crispr_ptmuts <- FilterForPointMuts(dataframe = crispr, outfile = "./../htscreens_giant_files/crispr_data_19Q1_g2pgenes_ptmuts.rds")
```

@@

```{r, eval=FALSE}
crispr_signif_gene <- WilcoxonTestDF(dataframe = crispr_data_g2p, grouping = "Gene", outfile = "./data_munging/rds/crispr_signif_g2p_gene.rds")
crispr_signif_lineage <- WilcoxonTestDF(dataframe = crispr_data_g2p, grouping = c("Gene", "Primary.Disease"), outfile = "./data_munging/rds/crispr_signif_g2p_lineage.rds")

@@

# ctrp_signif_gene <- WilcoxonTestDF(dataframe = ctrp_ptmuts, grouping = "Drug_Gene", outfile = "./data_munging/rds/ctrp_signif_g2p_gene.rds")
# ctrp_signif_gene$G2P_Drug_Gene <- ifelse(ctrp_signif_gene$Drug_Gene %in% g2p_druggenes, TRUE, FALSE)
# ctrp_signif_lineage <- WilcoxonTestDF(dataframe = filter(ctrp_ptmuts, Drug_Gene_Lin %in% g2p_dgls), grouping = "Drug_Gene_Lin", outfile = "./data_munging/rds/ctrp_signif_g2p_lineage.rds")

>>>>

    # Get all G2Pgene-cell line combinations
    grid <- expand.grid("Cell_line" = unique(filt$Cell_line), "Gene" = g2p_genes)
    data <-  merge(filt, grid, by = "Cell_line", all.y = TRUE)
    print("Merged grid.")
    # Merge metadata
    data <-  merge(data, select(ccl_meta, Cell_line, Histology, CCLE_Name, Broad_ID, Primary.Disease), by = c("Cell_line", "Histology", "CCLE_Name", "Broad_ID"), all.x = TRUE)
    ## Add columns for G2P drug-gene and drug-gene-lineage sorting
    data$Drug_Gene <- paste0(data$Drug, "_", data$Gene)
    data$G2P_Drug_Gene <- ifelse(data$Drug_Gene %in% g2p_druggenes, "Yes", "No")
    data$Drug_Gene_Lin <- paste0(data$Drug, "_", data$Gene, "_", data$Primary.Disease)
    data$G2P_Drug_Gene_Lin <- ifelse(data$Drug_Gene_Lin %in% g2p_dgls, "Yes", "No")
    print("Merged metadata.")
  }
  if(database == "CRISPR") {
    print(paste0("Processing ", database, " data."))
    # Melt CRISPR dataset for merging
    data <- melt(crispr_all, id.vars = "Broad_ID", measure.vars = colnames(crispr_all)[2:ncol(crispr_all)], variable.name = "Gene", value.name = "Score")
    # Merge cell line metadata
    data <- merge(data, select(ccl_meta, Cell_line, Histology, CCLE_Name, Broad_ID, Primary.Disease), by = "Broad_ID", all.x = TRUE)
    data$Drug_Gene <- paste0(data$Drug, "_", data$Gene)
    data$Drug_Gene_Lin <- paste0(data$Drug, "_", data$Gene, "_", data$Primary.Disease)
    print("Merged metadata.")
  }

@@

ccle <- MergeOmicsG2P(database = "CCLE", outfile = "./data_munging/rds/ccle_data_19Q1_g2pgenes.rds")
ctrp <- MergeOmicsG2P(database = "CTRP", outfile = "./../htscreens_giant_files/ctrp_data_19Q1_g2pgenes.rds")
gdsc <- MergeOmicsG2P(database = "GDSC", outfile = "./data_munging/rds/gdsc_data_19Q1_g2pgenes.rds")

ccle_ptmuts <- FilterForPointMuts(dataframe = ccle, outfile = "./data_munging/rds/ccle_data_19Q1_g2pgenes_ptmuts.rds")
ctrp_ptmuts <- FilterForPointMuts(dataframe = ctrp, outfile = "./../htscreens_giant_files/ctrp_data_19Q1_g2pgenes_ptmuts.rds")
gdsc_ptmuts <- FilterForPointMuts(dataframe = gdsc, outfile = "./data_munging/rds/gdsc_data_19Q1_g2pgenes_ptmuts.rds")
```

@@

```{r, eval=FALSE}
ccle <- readRDS("./data_munging/rds/ccle_data_19Q1_g2pgenes.rds")
ctrp <- readRDS("./../htscreens_giant_files/ctrp_data_19Q1_g2pgenes.rds")
gdsc <- readRDS("./data_munging/rds/gdsc_data_19Q1_g2pgenes.rds")
ccle_ptmuts <- readRDS("./data_munging/rds/ccle_data_19Q1_g2pgenes_ptmuts.rds")
ctrp_ptmuts <- readRDS("./../htscreens_giant_files/ctrp_data_19Q1_g2pgenes_ptmuts.rds")
gdsc_ptmuts <- readRDS("./data_munging/rds/gdsc_data_19Q1_g2pgenes_ptmuts.rds")

# Testing
test_ccle <- ccle_ptmuts %>% count(Gene, Mutation_Status)
test_ctrp <- ctrp_ptmuts %>% count(Gene, Mutation_Status)
test_gdsc <- gdsc_ptmuts %>% count(Gene, Mutation_Status)
```

@@

```{r, eval=TRUE}
crispr_data <- readRDS("./../htscreens_giant_files/crispr_data_19Q1.rds")

# Filter G2P CRISPR for point mutations
crispr_ptmuts <- filter(crispr_data, PointMutation == 1)
crispr_ptmuts <- subset(crispr_ptmuts, select = -c(PointMutation))
# Filter CRISPR for G2P genes in G2P
crispr_g2p <- filter(crispr_ptmuts, Gene %in% g2p_genes)
crispr_g2p <- unique(crispr_g2p)
crispr_g2p$Gene <- factor(crispr_g2p$Gene)
crispr_g2p$Mutation_Status <- factor(crispr_g2p$Mutation_Status)
crispr_g2p$Primary.Disease <- factor(crispr_g2p$Primary.Disease)
```

Test merging and filtering functions on CRISPR data:
```{r, eval=FALSE}
# crispr_all <- read.delim("./data_munging/19Q1/gene_effect_corrected_20190305.csv.gz", sep = ",", header  = TRUE, check.names = FALSE)
# # Remove Entrez gene IDs from colnames
# colnames(crispr_all) <- gsub(" .*", "", colnames(crispr_all))
# colnames(crispr_all)[1] <- "Broad_ID"

crispr_fxn <- MergeOmicsG2P(database = "CRISPR", outfile = "./../htscreens_giant_files/crispr_data_19Q1_g2pgenes_fxn.rds")
crispr_ptmuts_fxn <- FilterForPointMuts(dataframe = crispr_fxn, outfile = "./../htscreens_giant_files/crispr_data_19Q1_g2pgenes_ptmuts_fxn.rds")
```

@@

```{r, eval=TRUE}
crispr_signif_gene <- WilcoxonTestDF(dataframe = crispr_data_g2p, grouping = "Gene", outfile = "./data_munging/rds/crispr_signif_g2p_gene.rds")
crispr_signif_lineage <- WilcoxonTestDF(dataframe = crispr_data_g2p, grouping = c("Gene", "Primary.Disease"), outfile = "./data_munging/rds/crispr_signif_g2p_lineage.rds")

@@

ctrp_signif_gene <- WilcoxonTestDF(dataframe = ctrp_ptmuts, grouping = "Drug_Gene", outfile = "./data_munging/rds/ctrp_signif_g2p_gene.rds")
ctrp_signif_gene$G2P_Drug_Gene <- ifelse(ctrp_signif_gene$Drug_Gene %in% g2p_druggenes, TRUE, FALSE)
ctrp_signif_lineage <- WilcoxonTestDF(dataframe = filter(ctrp_ptmuts, Drug_Gene_Lin %in% g2p_dgls), grouping = "Drug_Gene_Lin", outfile = "./data_munging/rds/ctrp_signif_g2p_lineage.rds")
